Verification of Stencil Computations with Spatial Specifications

### 2016-11-13 paper.tex
equations and inequations

The equation (\textsc{$\ast$Dim}) uses syntactic sugar
where $R^d$ and $S^d$ denote regions whose dimension is $d$.
Thus, the rule shows that a product of two regions on 
the same dimension is equal to their addition.

@@

\begin{figure}[t]
\begin{align*}
\begin{array}{c}
%%
\trule{up}\dfrac{\forall u \in M \, \exists v \in N .  \, u \preceq v}
      {\cons{\textsf{upper}(N)}{M}}
\qquad
\trule{low}\dfrac{\forall v \in N \, \exists u \in M . \, u \preceq v}
      {\cons{\textsf{lower}(N)}{M}} \\[1.5em]
%%
\trule{exact}\dfrac{\cons{\textsf{lower}(N)}{M} \qquad \cons{\textsf{upper}(N)}{M}}
  {\cons{\textsf{exact}(N)}{M}} \\[1.5em]
%%
\trule{both}\dfrac{\cons{\textsf{lower}(M_1)}{M} \qquad \cons{\textsf{upper}(M_2)}{M}}
  {\cons{\textsf{both}(M_1, M_2)}{M}}\\[1.5em]
%%
\trule{$1$}
\dfrac{\cons{N}{M}}{\consA{\textsf{once}(N)}{\textsf{once}(M)}}
\;\;
\trule{$*$}
\dfrac{\cons{N}{M}}{\consA{\textsf{mult}(N)}{\textsf{mult}(M)}}
  \\[1.5em]
\trule{$1 < *$}
\dfrac{\cons{N}{M}}{\consA{\textsf{mult}(N)}{\textsf{once}(M)}}
\end{array}
\end{align*}
\caption{Consistency between models of code and specification}
\label{fig:consistency}
\vspace{-0.9em}
\end{figure}

>>>>

\paragraph{Interaction between \term{+} and \term{*}}

The last seven equations (from (\textsc{\term{*}Dim}) inclusive), explain
the algebraic behaviour of \term{+} and \term{*}.  Together, we see
that \term{+} is an associative, commutative, idempotent binary
operator that distributes with \term{*}, which is associative and
commutative. This distribution of \term{*} over \term{+} is key as it
is used internally to give a normal form for stencil specifications
akin to Disjunctive Normal Form (DNF) (where \term{+} is 
disjunction and \term{*} is conjunction). We revisit this
normalisation in \S~\ref{sec:inf-step4}.

The equation (\textsc{$\ast$Dim}) uses syntactic sugar
where $R^d$ and $S^d$ denote regions whose dimension is $d$.
This equation shows that a product of two regions on 
the same dimension is equal to their addition.

The equational theory provides not only an axiomatic semantics
for the language, which the user can use to rewrite their
specification, but it is also used extensively in the inference
procedure (\S~\ref{subsec:inference}) to simplify specifications.

\begin{theorem}[Soundness of equational theory]
\begin{equation*}
\forall S, T, n . \quad
S \equiv T \; \Rightarrow \;
\interp{S}_n = \interp{T}_n
\end{equation*}
\end{theorem}

\subsection{Approximation}

Figure~\ref{fig:inequations} defines a notion of sub-specifications
via the relation $<:^r$ on \textit{region}
syntax and then $<:$ on \textit{spec} syntax inductively.
Both relations are reflexive and transitive (we omit
these equations) and congruent
with respect to \term{+}, \term{*}, \term{atLeast}, \term{atMost},
and \term{readOnce}. The (\trule{\textsc{eq}}) rule connects region
equations with inequations. The (\trule{\textsc{over}}) rule explains
the notion of spatial over-approximation via overlapping regions.
The (\trule{\textsc{rep}}) allows the modifier that
enforces that every index is read just once to be dropped
as an approximation; (\trule{\textsc{bound}}) gives
a connection between the \term{atLeast} and \term{atMost}
approximation modifiers.

\begin{theorem}[Soundness of approximation]
\begin{equation*}
\forall S, T, M, n . \quad
S <: T \; \wedge (\consA{\interp{S}_n}{M})
\Rightarrow (\consA{\interp{T}_n}{M})
\end{equation*}
\end{theorem}

\begin{figure}[t]
\begin{align*}
\hspace{-0.5em}
\begin{array}{c}
%%
%% INEQUATIONS ON REGION
%% 
%\framebox{$\textit{region} <: \textit{region}'$} \\[0.85em]
%
%\dfrac{}{R <:^r R}\trule{\textsc{refl}} \qquad \dfrac{P <:^r R \quad R <:^r Q}{P <:^r
%  Q}\trule{\textsc{trans}} 
%\qquad
%\\[1.5em]
\vspace{-0.4em}
\setlength{\arraycolsep}{0.1em}
\dfrac{\hspace{3em} m \leq n \qquad\qquad r \implies s \hspace{3em}}
{\begin{array}{rl}
\stenFwdSR{m}{d}{r} & <:^r \stenFwdSR{n}{d}{s}
%& \;\, ds \subseteq es \wedge n \leq m
\\
\wedge \; \stenBwdSR{m}{d}{r} & <:^r \stenBwdSR{n}{d}{s}
%& \;\, ds \subseteq es \wedge n \leq m
\\
\wedge \; \stenCenSR{m}{d}{r} & <:^r \stenCenSR{n}{d}{s}
%& \;\, ds \subseteq es \wedge n \leq m
\\
\wedge \; \stenBwdSR{m}{d}{r} & <:^r \stenCenSR{n}{d}{s} \\
\wedge \; \stenFwdSR{m}{d}{r} & <:^r \stenCenSR{n}{d}{s} \\[1.2em]
\end{array}}\trule{\textsc{over}} \\[-0.2em]
\hspace{-0.5em}
\dfrac{R_1 <:^r S_1 \quad R_2 <:^r S_2}
      {R_1 \, \texttt{+} \, R_2 <:^r S_1 \, \texttt{+} \, S_2}
\trule{\textsc{cong}\texttt{+}} 
\dfrac{R_1 <:^r S_1 \quad R_2 <:^r S_2}
      {R_1 \, \texttt{*} \, R_2 <:^r S_1 \, \texttt{*} \, S_2}
\trule{\textsc{cong}\texttt{*}} \\[1.5em]
%%
%% INEQUATIONS ON SPEC
%%
%\framebox{$\textit{spec} <: \textit{spec}'$} \\[0.8em]
%
%
\dfrac{R <:^r S}{\term{atLeast}  S <: \term{atLeast} 
  R}\trule{\texttt{L}}
\;\;
\dfrac{R <:^r S}{\term{atMost}  R <: \term{atMost} 
  S}\trule{\texttt{M}} \\[1.5em]
\dfrac{R <:^r S}{R <: \term{atMost} \, S}\trule{\textsc{shrink}} \quad
\dfrac{R <:^r S}{S <: \term{atLeast} \, R}\trule{\textsc{grow}} \quad
 \\[1.5em]
%%
\dfrac{\mathsf{dims}(R) \cap \mathsf{dims}(S) = \emptyset}{R \ast S <: R}
\trule{\textsc{Gen$\ast$}}
\quad
\dfrac{R \equiv S}{R <: S}\trule{\textsc{eq}} 
%%
\\[1.5em]
\dfrac{R <: S}{\term{readOnce} \, R <: \term{readOnce} \,
  S}\trule{\textsc{cong}\texttt{R}}
\;\;
\dfrac{}{\term{readOnce} \, R <: R}\trule{\textsc{rep}}
%\dfrac{}{R <: \term{atMost} R} 
%\quad
%\dfrac{}{R <: \term{atLeast} R}\\[1.5em]
\end{array}
\end{align*}
\caption{Inequations $<:^r$ on \textit{regions} and
$<:$ on \textit{specs} (omitting reflexivity
and transitivity). }
\label{fig:inequations}
\vspace{-0.8em}
\end{figure}

@@

\begin{figure}[t]
\begin{align*}
\begin{array}{c}
%%
\trule{up}\dfrac{\forall u \in M, \, \exists v \in N .  \, u \preceq v}
      {\cons{\textsf{upper}(N)}{M}}
\qquad
\trule{low}\dfrac{\forall v \in Nm \, \exists u \in M . \, u \preceq v}
      {\cons{\textsf{lower}(N)}{M}} \\[1.5em]
%%
\trule{exact}\dfrac{\cons{\textsf{lower}(N)}{M} \qquad \cons{\textsf{upper}(N)}{M}}
  {\cons{\textsf{exact}(N)}{M}} \\[1.5em]
%%
\trule{both}\dfrac{\cons{\textsf{lower}(M_1)}{M} \qquad \cons{\textsf{upper}(M_2)}{M}}
  {\cons{\textsf{both}(M_1, M_2)}{M}}\\[1.5em]
%%
\trule{$1$}
\dfrac{\cons{N}{M}}{\consA{\textsf{once}(N)}{\textsf{once}(M)}}
\;\;
\trule{$*$}
\dfrac{\cons{N}{M}}{\consA{\textsf{mult}(N)}{\textsf{mult}(M)}}
  \\[1.5em]
\trule{$1 < *$}
\dfrac{\cons{N}{M}}{\consA{\textsf{mult}(N)}{\textsf{once}(M)}}
\end{array}
\end{align*}
\caption{Consistency between models of code and specification}
\label{fig:consistency}
\vspace{-0.9em}
\end{figure}
